---
title: "Step1_Prepare_data"
author: "Liam D. Bailey"
date: "18/10/2019"
output:
  pdf_document: default
  html_document: default
---

#Overview:

In this vignette, we use the functions `estimate_sensitivity()` and `estimate_exposure()` to determine the phenological sensitivity and climate exposure of all 67 population/species combinations.

**Note:** Although we include code for running temperature windows and structural equation models this is not executed in the vignette because it takes a long time.

**Note:** Mean daily temperature data from ECA&D v17.0 is not stored with the package, but can be requested from ECA&D (https://www.ecad.eu/).

```{r, message=FALSE}

#Load packages and set options
options(scipen = 200)
library(tidyverse)
#climwin
library(climwin)
#Load internal functions
library(bailey2021)

```

# Run climwin and SEM

- We extract mean daily temperature data and load laying date data, which is stored in the package files.

- We run a sliding window analysis for all populations.

- We run a randomisation procedure to overcome issues of multiple test.

- We run SEM for all populations with 1,000 bootstrap iterations to determine coefficients.

```{r, eval = FALSE}

#Load mean temperature data
#Temperature data file can be requested from https://www.ecad.eu/ archive
temp_data  <- raster::stack(here::here("data/unshared_files/tg_0.25deg_reg_v17.0.nc"))

#Load laying date data
data("Lay_date_summary")

#Run climwin and randwin with 100 iterations.
#Run SEM with 1000 bootstrap iterations to determine SEs
#after accounting for possible shared trends
all_results <- estimate_sensitivity(input_data = Lay_date_summary, temp_data = temp_data,
                                    randwin = TRUE, repeats = 100, SEM = TRUE, i = 1000)

```

# Create data frame with results

Once we have run climwin, randwin, and SEM we can then determine the PAICc significance of climate windows in every population.

```{r, warning=FALSE}

#Link to files output from climwin
all_files <- list.files(here::here("./data/unshared_files"), pattern = "rand.RDS", full.names = TRUE)

pb <- txtProgressBar(max = length(all_files), style = 3)

sem_output <- purrr::map_dfr(.x = 1:length(all_files),
                             .f = ~{
                               
                               setTxtProgressBar(pb, value = ..1)
                               
                               SEM_path <- gsub(all_files[..1], pattern = "_rand.RDS", replacement = "_SEM.RDS")
                               climwin_path <- gsub(all_files[..1], pattern = "_rand.RDS", replacement = ".RDS")
                               randwin_path <- all_files[..1]
                               
                               #Firstly, extract the output from run_SEM that has the overview of the top model
                               output <- readRDS(SEM_path) %>% 
                                 dplyr::mutate(PAIC = as.character(climwin::pvalue(
                                   dataset = readRDS(climwin_path)[[1]]$Dataset,
                                   datasetrand = readRDS(randwin_path)[[1]],
                                   metric = "AIC")))
                               
                               output$PAIC_dbl <- ifelse(output$PAIC == "<0.001", 0.001, as.numeric(output$PAIC))
                               
                               return(output)
                               
                             })

```

# Determine rate of climate change

Then we use `estimate_exposure` to determine the rate of climate change over time between 1950 and 2017. Save all this data in a single file for model fitting.

```{r}

climate_sensitivity_and_exposure <- estimate_exposure(sem_output)

```

We now have a data frame that includes:

- The best window estimated by climwin.

- The significance of this window, using PAICc method with randwin and 100 iterations.

- The coefficient estimate for the relationship between temperature and mean annual laying date
after detrending with SEM using bootstrapping with 1,000 iterations (i.e. sensitivity)

- The rate of climate change at each site during the population specific temperature window over the period 1950 - 2017
(i.e. exposure).

Next we can join in the physical characteristics of each population that we use for analysis. This includes:

- Habitat type
- Coordinates

```{r}

data("Pop_info")

climate_sensitivity_and_exposure <- climate_sensitivity_and_exposure %>% 
  dplyr::left_join(Pop_info, by = "Pop_ID")

```

We can now save this file in the package for use in analyses.

```{r}

usethis::use_data(climate_sensitivity_and_exposure, overwrite = TRUE)

```
